Contextuality and Nonlocality in Discrete Quantum Theory

### 2016-02-08 DQC3.tex
"6.1 Gleason's Theorem in DQT" Start editing...

Recall the measurement process discussed in section \ref{sec:Critical-Analysis-CQT}.
Given a physical system represented by a state $\ket{\Psi}$, and
an observable ${\mathbf{O}}$, whose eigenspectrum is given by ${\mathbf{O}}\left|i\right\rangle =\lambda_{i}\left|i\right\rangle $.
Then the Born rule of CQT tells us that the probability of obtaining
the eigenvalue $\lambda_{i}$ as an outcome of the measurement is
given by 
\begin{eqnarray}
\begin{array}{rcccl}
{\cal P}_{\Psi}\left(\ket{i}\right) & = & {\cal P}_{\Psi}\left(\lambda_{i}\right) & = & \frac{\ip{\Psi}{i}\ip{i}{\Psi}}{\ip{\Psi}{\Psi}\ip{i}{i}}\end{array}\ ,\label{prob1.eq-6}
\end{eqnarray}
If the system is a mixed state represented by a density matrix 
\[
\rho=\sum_{j}p_{j}\frac{\proj{\Psi_{j}}}{\ip{\Psi_{j}}{\Psi_{j}}}\textrm{ ,}
\]
we can extend the Born rule (\ref{prob1.eq-6}) and get 
\begin{eqnarray}
{\cal P}_{\rho}\left(\ket{i}\right)={\cal P}_{\rho}\left(\lambda_{i}\right) & = & \sum_{j}p_{j}{\cal P}_{\Psi_{j}}\left(\ket{i}\right)\label{prob1.eq-2}
\end{eqnarray}

Let $\C^{*}=\C\backslash\left\{ 0\right\} $ and $\C^{d*}=\left\{ \ket{\Psi}\in\C^{d}\mid\ip{\Psi}{\Psi}\ne0\right\} $
be the set of non-zero vectors. Because normalization and global phase
gives the same state, and ${\cal P}_{\rho}:\C^{d*}\rightarrow\left[0,1\right]$
should be a state-function, we should have 
\begin{equation}
\textrm{For any vector }\ket{i}\in\C^{d*}\textrm{, and }\gamma\in\C^{*}\textrm{ , }{\cal P}_{\rho}(\ket{i})={\cal P}_{\rho}(\gamma\ket{i})\textrm{ .}\label{eq:parallel-DQT(I)}
\end{equation}
More importantly, ${\cal P}_{\rho}$ defines a probability which means
\begin{equation}
\textrm{For any orthonormal basis }\{\ket{i}\}\textrm{ , }{\displaystyle \sum_{i=0}^{d-1}{\cal P}_{\rho}\left(\ket{i}\right)=1}\textrm{ .}\label{eq:total-probability}
\end{equation}
Conversely, if a state-function ${\cal P}:\C^{d*}\rightarrow\left[0,1\right]$
is a probability, we can identify ${\cal P}$ to a mixed state $\rho$
by Gleason's theorem.
\begin{thm}
(Gleason's Theorem)

For $d\ge3$, for any ${\cal P}:\C^{d*}\rightarrow\left[0,1\right]$,
if ${\cal P}$ is a state-function and probability, i.e.,
\begin{eqnarray}
 &  & \textrm{For any vector }\ket{i}\in\C^{d*}\textrm{, and }\gamma\in\C^{*}\textrm{ , }{\cal P}(\ket{i})={\cal P}(\gamma\ket{i})\textrm{ ,}\label{eq:parallel-DQT(I)-1}\\
 &  & \textrm{For any orthonormal basis }\{\ket{i}\}\textrm{ , }{\displaystyle \sum_{i=0}^{d-1}{\cal P}\left(\ket{i}\right)=1}\textrm{ ,}\label{eq:total-probability-1}
\end{eqnarray}
there is a density matrix $\rho$ such that ${\cal P}={\cal P}_{\rho}$~\cite{Redhead1987-REDINA,gleason1957,peres1995quantum}.\footnote{In order to simplify the statement, we consider consider the probability
on the state vectors instead of the projectors or closed subspaces
as in the Gleason's original paper.}
\end{thm}
In contrast, when $d=2$, consider ${\cal P}:\C^{2*}\rightarrow\left[0,1\right]$
such that if we restrict ${\cal P}$ on the real unit circle, we will
have a discontinuous function 
\[
{\cal P}
\left(\begin{array}{c} \cos\theta\\ \sin\theta \end{array}\right)=
\cases{ 
0 & , if $0 \le \theta < \frac{\pi}{2}$\\
1 & , if $\frac{\pi}{2} \le \theta < \pi$ } 
\]
${\cal P}$ satisfies equation (\ref{eq:parallel-DQT(I)-1}) and (\ref{eq:total-probability-1})
because we have 
\[
{\cal P}\left(\begin{array}{c}
\cos\theta\\
\sin\theta
\end{array}\right)+{\cal P}\left(\begin{array}{c}
-\sin\theta\\
\cos\theta
\end{array}\right)={\cal P}\left(\begin{array}{c}
\cos\theta\\
\sin\theta
\end{array}\right)+{\cal P}\left(\begin{array}{c}
\cos\left(\theta+\frac{\pi}{2}\right)\\
\sin\left(\theta+\frac{\pi}{2}\right)
\end{array}\right)=1\textrm{ .}
\]
However, ${\cal P}_{\rho}$ is always continuous. Therefore, we can
never find a density matrix $\rho$ such that ${\cal P}={\cal P}_{\rho}$.

>>>>

Recall the measurement process discussed in section \ref{sec:Critical-Analysis-CQT}.
Given an observable ${\mathbf{O}}$ whose eigenspectrum is given by
${\mathbf{O}}\left|i\right\rangle =\lambda_{i}\left|i\right\rangle $,
when we measure ${\mathbf{O}}$ on a system in a state $\ket{\Psi}$,
the probability is given by the Born rule of CQT 
\begin{eqnarray}
\begin{array}{rcccl}
{\cal P}_{\Psi}\left(\ket{i}\right) & = & {\cal P}_{\Psi}\left(\lambda_{i}\right) & = & \frac{\ip{\Psi}{i}\ip{i}{\Psi}}{\ip{\Psi}{\Psi}\ip{i}{i}}\end{array}\ ,\label{prob1.eq-6}
\end{eqnarray}
If the system is in a mixed state represented by a density matrix
\[
\rho=\sum_{j}p_{j}\frac{\proj{\Psi_{j}}}{\ip{\Psi_{j}}{\Psi_{j}}}\textrm{ ,}
\]
the Born rule (\ref{prob1.eq-6}) can be extended to 
\begin{eqnarray}
{\cal P}_{\rho}\left(\ket{i}\right)={\cal P}_{\rho}\left(\lambda_{i}\right) & = & \sum_{j}p_{j}{\cal P}_{\Psi_{j}}\left(\ket{i}\right)\textrm{ ,}\label{prob1.eq-2}
\end{eqnarray}

There are two important properties ${\cal P}_{\rho}:\C^{d*}\rightarrow\left[0,1\right]$
should have. First, if we normalize a vector or multiply it by a constant
factor, it still represents the same state, and ${\cal P}_{\rho}$
is a function of state. Thus, we have ${\cal P}_{\rho}(\ket{i})={\cal P}_{\rho}(\gamma\ket{i})$
for any $\gamma\in\C^{*}$ and $\ket{i}\in\C^{d*}$, where $\C^{*}=\C\backslash\left\{ 0\right\} $
and $\C^{d*}=\left\{ \ket{\Psi}\in\C^{d}\mid\ip{\Psi}{\Psi}\ne0\right\} $
is the set of non-zero vectors. Second, ${\cal P}_{\rho}$ defines
a probability, and the total probability should be one, i.e., ${\displaystyle \sum_{i=0}^{d-1}{\cal P}_{\rho}\left(\ket{i}\right)=1}$
for any orthonormal basis $\left\{ \ket{i}\right\} $.

Conversely, if ${\cal P}:\C^{d*}\rightarrow\left[0,1\right]$ is a
function of state and defines a probability, ${\cal P}$ can be identified
to a mixed state $\rho$ by Gleason's theorem.
\begin{thm}[Gleason's]
For $d\ge3$, if ${\cal P}:\C^{d*}\rightarrow\left[0,1\right]$ is
a function of state and defines a probability, i.e.,
\begin{eqnarray}
 &  & {\cal P}(\ket{i})={\cal P}(\gamma\ket{i})\textrm{ for any vector }\ket{i}\in\C^{d*}\textrm{, and }\gamma\in\C^{*}\textrm{ ,}\label{eq:parallel-DQT(I)-1}\\
 &  & {\displaystyle \sum_{i=0}^{d-1}{\cal P}\left(\ket{i}\right)=1}\textrm{ for any orthonormal basis }\{\ket{i}\}\textrm{ ,}\label{eq:total-probability-1}
\end{eqnarray}
there is a density matrix $\rho$ such that ${\cal P}={\cal P}_{\rho}$~\cite{Redhead1987-REDINA,gleason1957,peres1995quantum}.\footnote{In order to simplify the statement, we consider the probability on
the state vectors instead of the projectors or closed subspaces as
in the Gleason's original paper.}
\end{thm}
Gleason's theorem requires $d\ge3$ because different orthonormal
bases may share their vectors. For example, $\left\{ \ket{0},\ket{1},\ket{2}\right\} $
and $\left\{ \ket{0},\frac{1}{\sqrt{2}}\left(\ket{1}+\ket{2}\right),\frac{1}{\sqrt{2}}\left(\ket{1}-\ket{2}\right)\right\} $
share the vector $\ket{0}$. This implies ${\cal P}$ should be continuous
and further proves the theorem. In contrast, when $d=2$, there is
only one state orthogonal to a given state. Hence, every equation
(\ref{eq:total-probability-1}) is either identical or share no common
${\cal P}\left(\ket{i}\right)$. Therefore, when we solve the system
of linear equation (\ref{eq:total-probability-1}), we can actually
solve them individually and get a discontinuous function. For example,
${\cal P}:\C^{2*}\rightarrow\left[0,1\right]$ such that if we restrict
${\cal P}$ on the real unit circle 
\[
{\cal P}
\left(\begin{array}{c} \cos\theta\\ \sin\theta \end{array}\right)=
\cases{ 
0 & , if $0 \le \theta < \frac{\pi}{2}$\\
1 & , if $\frac{\pi}{2} \le \theta < \pi$ } 
\]
${\cal P}$ satisfies equation (\ref{eq:parallel-DQT(I)-1}) and (\ref{eq:total-probability-1})
because we have 
\[
{\cal P}\left(\begin{array}{c}
\cos\theta\\
\sin\theta
\end{array}\right)+{\cal P}\left(\begin{array}{c}
-\sin\theta\\
\cos\theta
\end{array}\right)={\cal P}\left(\begin{array}{c}
\cos\theta\\
\sin\theta
\end{array}\right)+{\cal P}\left(\begin{array}{c}
\cos\left(\theta+\frac{\pi}{2}\right)\\
\sin\left(\theta+\frac{\pi}{2}\right)
\end{array}\right)=1\textrm{ .}
\]
However, ${\cal P}_{\rho}$ is always continuous. Therefore, we can
never find a density matrix $\rho$ such that ${\cal P}={\cal P}_{\rho}$.
